VFTS682: a confirmed dynamical ejection?

### 2018-04-30 VFTS682.tex

\defcitealias{tauris:98}{TT98}
\defcitealias{ramirez-agudelo:15}{R-A15}

@@

\title{Massive Runaway and Walkaway Stars}
\subtitle{A study of the kinematical imprints of the physical processes \\ governing the evolution and explosion of their binary progenitors}

\author{M.~Renzo\inst{1} \and E.~Zapartas\inst{1} \and S.~E.~de~Mink\inst{1} \and
 Y.~G\"otberg\inst{1} \and S.~Justham\inst{2,3} \and
 R.~J.~Farmer\inst{1} \and R.~G.~Izzard\inst{4}
 \and S.~Toonen\inst{1}
 \and H.~Sana\inst{5}.}

\institute{{Astronomical Institute Anton Pannekoek, University of
    Amsterdam, 1098 XH Amsterdam, The Netherlands}
  \and{School of Astronomy \& Space Science, University of the Chinese
    Academy of Sciences, Beijing 100012, China}
  \and{National Astronomical Observatories, Chinese Academy of
    Sciences, Beijing 100012, China}
  \and{Astrophysics Research Group, Faculty of Engineering and Physical Sciences, University of Surrey, Guildford, Surrey, GU2 7XH, United Kingdom}
  \and{Institute of Astronomy, KU Leuven, Celestijnenlaan 200 D, B-3001 Leuven, Belgium}}

\offprints{M.~Renzo, \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}}
\date{}
\abstract{Anticipating the kinematic constraints from
  the Gaia mission, we perform an extensive numerical study of
  the evolution of massive binary systems to predict the peculiar
  velocities that stars obtain when their companion collapses and disrupts the system. Our aim is to (1) identify which predictions are robust against model uncertainties and assess their implications, (2) investigate which physical processes leave a clear imprint and may therefore be constrained observationally and (3) provide a suite of publicly available model predictions.  
We find that $22_{-8}^{+26}$\% of all massive binary systems merge
prior to the first core collapse in the system. Of the remainder,
$86_{-9}^{+11}\%$ become unbound because of the
core-collapse. Remarkably, this rarely produce runaway stars (i.e.,
stars with velocities above $30\,\kms$). These are outnumbered by more
than an order of magnitude by slower unbound companions, or ``walkaway
stars’’. This is a robust outcome of our simulations and is due to the
reversal of the mass ratio prior to the explosion and widening of the
orbit, as we show analytically and numerically. We estimate a
$10^{+5}_{-8}$\% of massive stars to be walkaways and only $0.5^{+1.0}_{-0.4}$\% to be runaways, nearly all of which have
accreted mass from their companion. Our findings are consistent with
earlier studies, however the low runaway fraction we find is in
tension with observed fractions $\sim$10\%. If Gaia confirms these high fractions of
massive runaway stars resulting from binaries, it would imply that we
are currently missing physics in the binary models. Finally, we show
that high end of the mass distributions of runaway stars is very
sensitive to the assumed black hole natal kicks and propose this as a
potentially stringent test for the explosion mechanism. We also
discuss companions remaining bound which can evolve into X-ray
 and gravitational wave sources.}

\keywords{stars: kinematics, binaries, massive, supernovae }
\maketitle{}

@@

Stars with initial mass larger than
about $7.5\,M_\odot$ are the progenitors of black holes (BH) and
neutron stars (NS). These stars play an important role in shaping galaxies
through their radiative, chemical, mechanical feedback
(e.g., \citealt{larson:74, Heger+2003, ceverino:09, lamers:13}, see
also \citealt{langer:12, nomoto:13} for reviews).
Most young, unevolved, massive stars have a nearby companion with
which they form a close binary system
\citep[e.g.,][]{kobulnicky:07,mason:09,sana:11,sana:12,kiminki:12,chini:12,kobulnicky:14,almeida:17,demarco:17}. 
Binary systems that remain bound throughout the entire evolution of both
stars can give rise to many exotic phenomena, including X-ray binaries
\citep[e.g., ][]{gott:71,Bolton1972, Webster+1972,
  van-den-Heuvel+1972}, binary neutron stars
\citep[e.g.,][]{hulse:75,wijers:92}, gamma-ray bursts
\citep[e.g.,][]{izzard:04b,becerra:16, LVC:17:grb}, and gravitational wave events
\citep{LVC:16b, BNSmerger:17}.  
However, only a small fraction of massive stars that are born as
binary systems are expected to stay together their entire lives. The
majority of systems is disrupted by the first core collapse event,
which can separate the newly formed compact object from its former
companion star
\citep[][]{zwicky:57,blaauw:61, tauris:98, belczynski:99,
  belczynski:08, eldridge:11}.

Single NSs can be detected as pulsars or magnetars. Many of
them are observed to have large proper motions (e.g.,
\citealt{gunn:70,gott:70,lyne:94,kaspi:96,hobbs:05,chatterjee:05}, but see also \citealt{verbunt:17}).
Conversely, single BHs are only detectable under special
circumstances, for example through lensing events when passing in
front of a background star \citep[e.g.,][]{wyrzykowski:16}, or if they
accrete gas from the ambient medium
\citep[e.g.,][]{fender:13, gaggero:17}. Therefore, to probe the
population of stellar-mass black holes we are limited to X-ray or
gravitational wave observations that can only target the rare
cases that remain bound to their companion and are close enough for
the BH to accrete or merge. To learn more about the black holes that
form in less special cases, it is worth to investigate the imprints
they may leave on their former companion star.

The primary focus of this study is the population of unbound main sequence companions,
which can be identified observationally because of their peculiar
spatial velocities compared to the surrounding population. Their velocities can either be detected as proper
motions (i.e.,\ their motion in the plane of the sky measured directly
from the displacement of the star) or as radial velocities (i.e.,\ their
motion perpendicular to the plane of the sky, measured from the
Doppler shift of the spectral lines). Large spatial velocities have
been inferred for a significant sub-population of young massive stars
\citep[e.g.][]{blaauw:61, Cruz-Gonzalez+1974, gies:86, gies:87,
  hoogerwerf:00, hoogerwerf:01, tetzlaff:11, boubert:18}.

\citet[][]{blaauw:61} introduced the term ``runaway stars'' for those %stars
in the fast tail of the velocity distribution for a given spectral type. The typical threshold
adopted to define the tail of this distribution for O and B-type stars is $v \gtrsim 30\,\mathrm{km \ s^{-1}}$
\citep[e.g.,][]{blaauw:56, gies:86, dedonder:97, hoogerwerf:00, hoogerwerf:01, dray:05, eldridge:11}, although sometimes other values
have been considered \citep[e.g., $40\,\mathrm{km\ s^{-1}}$
in][]{blaauw:61, dewit:05, boubert:18}. However, as we will argue based on simulations presented in this work, the majority of unbound companions is expected to exhibit velocities well below these thresholds.  
We will refer to these slow unbound former companions resulting from
disrupted binary systems as ``walkaway stars'' \citep[to our knowledge first coined
by][]{de-Mink+2012} to distinguish them from the faster counterparts.

A proposed explanation for the large spatial velocity of runaways is that
they originate from disrupted binary systems \citep[]{zwicky:57,
  blaauw:61, boersma:61},
which naturally explains the lower number of companions they have,
compared to typical massive stars \citep[e.g.,][]{blaauw:61, gies:86, sana:14}. \Figref{fig:cartoon} sketches the
typical evolution of such a binary system. Most binaries are disrupted
at the time of the first core-collapse.

An alternative mechanism to produce stars with peculiar spatial
velocities is dynamical ejection from a star cluster \citep[e.g.,][]{poveda:67, leonard:91, Evans+2010a,
  fujii:11, allison:12, oh:16}. Dynamical interactions with
a supermassive BH can also disrupt a binary, but the ejection
velocities achieved in this scenario are typically much higher
\citep[$\gtrsim$$10^3\,\mathrm{km\ s^{-1}}$,][]{hills:88}.

Both mechanisms, i.e. the ``binary
disruption scenario'' and the ``dynamical ejection scenario'', are
expected to act in nature, but their relative importance is not not well
constrained \citep[e.g.,][]{hoogerwerf:00,hoogerwerf:01,guseinov:05}.
\citet{hoogerwerf:01} analyzed the properties of a sample of 56 nearby
runaway stars and 9 radio pulsars, and traced back the runaways to
their most likely parent stellar group. From the sub-sample for which a
clear identification of the parent group was possible, they estimated that the disruption of binaries is
responsible for roughly two thirds of observed runaways \citep[see
also][]{gies:86, gies:87, stone:91, hoogerwerf:00, dincel:15,
  boubert:17b}. However, this claim could not be confirmed by
\cite{jilinski:10}.

Unbound stars resulting from the disruption of a binary are of
potential interest for several topical questions in
astrophysics. Their kinematics and stellar properties bear imprints of the uncertain physical processes that govern the evolution of their
binary progenitor systems, including the expected phase of mass
transfer between the two stars (phase B. in
\Figref{fig:cartoon}). Detailed knowledge of the spatial velocity of
runaway stars can improve the accuracy of wind mass-loss rate
determinations relying on their bow shocks \citep[e.g.][]{gull:79,
  kobulnicky:18}. Of particular interest is the
question whether they can provide any unique constraints on the physics of
core-collapse, in particular the natal kick on the compact objects
that they produce. Such kicks are expected either from asymmetries in
the explosion and/or the neutrino emission
\citep[][]{shklovskii:70,kusenko:96,fryer:04,socrates:05,wongwathanarat:13,
  janka:13,janka:17}, and they determine which systems disrupt and
eject the companion star and which systems remain bound and thus have a
potential as future X-ray and gravitational wave sources.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=0.5\textwidth]{./figures/cartoon}
  \caption{Schematic depiction of the evolution of a close massive
    binary through stable mass transfer. The
    evolution forks at the first core-collapse event in the system: the vast
  majority of the systems is disrupted and produces a runaway or
  walkaway star which can travel far from its birth location. The
  systems remaining bound are possible progenitors of X-ray binaries
  and gravitational wave sources. The fraction of binaries disrupted
  $\mathcal{D}$ comes from the simulations presented in this study.
}
  \label{fig:cartoon}
\end{figure}

These unbound stars have also been considered for their potential
importance as non-canonical sources of stellar feedback
\citep[e.g.,][]{larson:74, ceverino:09}. 
They can travel long distances and end their lives tens to hundreds of
parsecs away from their birth location. Because of their motion, their
ionizing photons are
less likely to be absorbed by their birth clouds. Therefore, in the context of the
re-ionization of the early universe, the ionizing radiation of stars
ejected from a binary are more likely to
escape and become available for ionization of hydrogen in the
intergalactic medium \citep[e.g.][]{conroy:12}. This has been
suggested to be of importance in the 
context of the epoch of reionization
\citep[e.g.,][]{conroy:12,kimm:14,Ma+2016}. Furthermore they are
expected to explode in lower density regions, which may enhance the
impact they have as sources of turbulence in the ISM \citep[e.g.][]{gatto:15}.

The present and upcoming second Gaia data releases provide an important
observational motivation for this study \citep[e.g.,][]{perryman:01}. The determination of
distances, proper motions and radial velocities by the Gaia satellite
is expected to drastically increase the available sample of massive
stars with precisely known velocities
\citep{maoileidigh:05,  cacciari:09, Jeffries+2014,
  Gaia-Collaboration+2016, Gaia-Collaboration+2016a}.

In this study, we present a systematic theoretical study of the
kinematical signatures of stars ejected from binary systems and how
these depend on the uncertain physical processes governing binary
stellar evolution. For this purpose we employ a rapid binary
population synthesis code, which we have updated to account for the main relevant processes that affect the disruption of the binary systems (see \Secref{sec:methods}).

We first provide insight in our simulations by presenting the results
for an individual system in \Secref{sec:example}.
We present in \Secref{sec:analytics} an analytic estimate of the
typical and maximum velocity of ejected companions expected from the
disruption of binaries. Then, in \Secref{sec:fiducial_res} we describe the results of a fiducial simulation of a
full population of binary stars, for which we adopt realistic
input assumptions provided by detailed simulations and/or observations
when available.  In addition, we present in
\Secref{sec:param_variation} an extensive grid of models where we vary the most relevant uncertain assumptions. This allows us to study (i) which predictions are robust against the model uncertainties and (ii) which uncertain physical processes leave a clear imprint on the observables that can be used to test and constrain them.

Various earlier studies have discussed the evolution and
interaction of populations of massive binary systems. Examples
include, but are not limited to, \citet[][]{vanbeveren:82, dedonder:97, Fryer+1997, Fryer+2001, fryer:12,
  Belczynski+2012b, repetto:12, Fragos+2013a, Grudzinska+2015,
  bray:18, boubert:18}. Most
of these studies focus on the minor fraction of systems
that remain bound and are the progenitors of X-ray binaries and/or
gravitational wave sources. A few studies focus specifically on
disrupted systems and unbound stars. Examples are
\cite{vanrensbergen:96,dedonder:97, dray:05, eldridge:11, bray:16, boubert:17b,
  zapartas:17b, boubert:18}. We expand on these studies by using updated physical
assumptions, and by focusing on which physical processes could be
constrained using the Galactic population of unbound stars. A
comparison to previously published results is given in \Secref{sec:literature}.

Our first main result is that the majority of disrupted binaries
ejects a slow-moving walkaway star ($v < 30\,\mathrm{km\ s^{-1}}$), rather than a faster runaway. This result is robust against
variations of uncertain parameters in the model, as we discuss in \Secref{sec:param_variation}.
This would imply that the more easily detected runaway stars only
reveal a small subset of the population of unbound stars that are
former companions of disrupted binary systems.  We discuss the implications in \Secref{sec:discussion}.

Our second main result is that the kinematic properties and absolute
number of unbound companions depend sensitively on assumptions
concerning the  natal kicks of BHs. We show the imprints this leaves on
the mass distribution and discuss whether future observations can be
used to constrain these processes.

Section~\ref{sec:bound} briefly describes the population of systems remaining bound after the first
core-collapse.

We also find that the fraction of runaways among massive stars
predicted by our simulations is much lower than the observed value in
all our parameter variations. This finding agrees with the results of
\cite{eldridge:11}, but is in potential contrast
with the observational results of \cite{hoogerwerf:01}, and needs further
investigation.

Finally, we discuss how our results  may be useful to
include the effect of unbound binaries in models for stellar
feedback. We will provide our numerical results upon publication. %provide our numerical results at \todo{where?}\todo{zenodo/CDS?}.

\section{Binary population synthesis calculations}
\label{sec:methods}

We carry out population synthesis calculations of isolated binaries with the rapid binary evolution code \texttt{binary\_c}
\citep[][]{izzard:04,izzard:06,izzard:09,demink:13,schneider:15,izzard:18}. This
code is based on the algorithms by \cite{tout:97,hurley:00,hurley:02}, which rely on the analytic fits to the
single stellar evolution models from \cite{pols:98}.

First, we compute a fiducial population using observationally
validated assumptions for the free parameters that describe the initial
conditions and physical assumptions.
Then, we check the robustness of our results (or
equivalently, to which assumptions they are most sensitive) by varying
the free parameters that are most relevant for the
velocity distribution of disrupted binaries. To limit
computational costs, we explore variations in each parameter
one-by-one while keeping the other parameters fixed to our
fiducial choices, following the approach of
\cite{Fragos+2013a,demink:13,demink:14,zapartas:17,zapartas:17b,
  belczynski:17}. Effectively, we treat
each parameter in our models as independent from the others, therefore this approach does not account for possible
correlations between either initial distributions \citep[e.g., the one between initial period and mass ratio
suggested by][]{moe:17} or uncertain
physical processes (e.g., core spin and natal kick amplitude). Varying
two (or more) parameters simultaneously, or changing the algorithmic
representation of the uncertain physical processes, might possibly result in
larger variations than those we present in
\Secref{sec:param_variation}.

\subsection{Initial distributions and parameters}

Each binary system in our calculations is characterized by a zero age
main sequence (ZAMS)
mass for the primary\footnote{Throughout this study, we define the primary star
  to be the initially more massive star, even
  if it becomes the less massive star in the binary during the evolution.}
$M_1^\mathrm{ZAMS}$, initial mass ratio $q\udef M_2/M_1$, and initial orbital
period $P^\mathrm{ZAMS}$. We build a grid in this parameter space and
weigh each system in our grid according to the initial distributions
described below.

We select $N_{M_1}$ primary stars with ZAMS mass $M_1^\mathrm{ZAMS}$
at logarithmically spaced intervals in the range $7.5\,M_\odot\leq M_1^\mathrm{ZAMS}
\leq 100\,M_\odot$. We weigh each primary star with an initial mass function (IMF) with
slope $\alpha=-2.3$ 
\citep[][]{kroupa:01}. In our model
variations, we explore values of $\alpha$ of -1.9
\citep[][]{schneider:18} and -3.

For each primary star of mass $M_1^\mathrm{ZAMS}$, we select $N_q$ secondaries with mass
$M_2^\mathrm{ZAMS} = qM_1^\mathrm{ZAMS}$, taken at regular intervals
in mass ratio $q$ between 0.1 and 1 assuming a
flat distribution \citep[slope $\kappa=0$, e.g.,][]{kouwenhoven:05,sana:12}. We consider in
\Secref{sec:param_variation} also variations with $\kappa=\pm1$.

Finally, for each pair of masses$(M_1^\mathrm{ZAMS},M_2^\mathrm{ZAMS})$ we choose $N_p$ different periods
$P^\mathrm{ZAMS}$ equally spaced in $\log_{10}(P^\mathrm{ZAMS}/\mathrm{days})$
between 0.15 and 5.5. We weigh the birth probability of each binary system with a
mass-dependent distribution for the initial orbital period: if $M_1<15\,M_\odot$ we assume
a flat distribution in $\log_{10}(P^\mathrm{ZAMS}/\mathrm{days})$
\citep[][]{opik:24,kobulnicky:07}; while for $M_1\geq15\,M_\odot$ we assume a
powerlaw distribution in $\log_{10}(P^\mathrm{ZAMS}/\mathrm{days})$ with exponent
$\pi=-0.55$, \citep[][]{sana:12}. We explore values of $\pi=0\,$and\,$\pi=-1$
(for all values of $M_1^\mathrm{ZAMS}$)
in our model variations. We chose our upper-limit on the initial period, $10^{5.5}$\,days, to include
wide systems in which the stars effectively evolve as single stars \citep[][]{demink:15}.

To limit the dimensions of the parameter space, we assume
all orbits to be initially circular (i.e., eccentricity
$e=0$). Because we focus on the population of stars resulting from the
disruption of post-interaction binaries, this assumption is not
critical \citep[][]{demink:15}: orbits are expected to
circularize
because of tides before or during mass transfer
(\citealt{belczynski:99,hurley:02}, but see also \citealt{eldridge:09}
for arguments against circularization of the orbit in post-main
sequence mass transfer).

In our fiducial model, we assume a canonical metallicity (i.e., mass fraction of
elements heavier than helium) $Z=0.02$. This value is slightly above
the most recent determination for solar neighborhood
\cite{asplund:09}, but it should generally describe young population
of massive stars in the Milky-Way. The models from \cite{pols:98}, which serve as
input for our computations, adopt the
isotopic mixture of \cite{anders:89}. In
\Secref{sec:param_variation}, we also consider $Z=0.0002$, $0.0047$, $0.008$,
and $0.03$.

We initialize the stellar rotation rate with a mass-dependent
equatorial velocity according to \cite{hurley:00}, assuming
alignment between the stellar spins and the orbital angular momentum. In one of our
parameter variations, we draw the initial spin velocity randomly from
the distribution given by \citealt{ramirez-agudelo:15} (R15, see last
line in \Tabref{tab:parameters}), but maintain the assumption of
alignment. However, changing the initial rotation rate has almost no
impact on our results because binary interaction processes overwrite it \citep[][]{demink:13}.

We build a grid in the initial parameter space, and initialize a
binary system in each cell of such grid. 
Each system is assigned a ``probability''
which can then be multiplied by the
amount of time spent by the system in that cell to obtain a quantity comparable to
observed number counts. See Appendix~\ref{app:data} for more details.

For each set of parameters, the total size of our model grid is
$N_{M_1}\times N_q \times N_p = 50 \times 50 \times 100 = 250\,000$ 
binary systems\footnote{This number is then multiplied by the number
  of natal kicks that we draw for each core-collapse, see \Secref{sec:SN_kick}.}.  Increasing this resolution to
$432\,000$ introduced no
significant variation in our results.

\subsection{Physical assumptions}

We follow the evolution of each binary system until the
first core-collapse (CC) event or until they merge. 
Stellar winds can have a dramatic effect on the evolution of single
stars \citep[e.g.,][]{renzo:17} and impact significantly the orbital
evolution in a binary. We include wind mass loss in our single star models as implemented by
\cite{demink:13}, i.e.\ we use a combination of mass loss rates from
\cite{vink:00,vink:01} for hot hydrogen-rich stars,
\cite{nieuwenhuijzen:90} for cool stars, % (or for luminosities larger than 4\,000$L_\odot$ if the mass loss rate
% is larger than what suggested by \cite{vink:00,vink:01})
 including a ``luminous blue variable'' enhancement as in
 \citealt{hurley:00}, and from \cite{hamann:98} reduced by a factor of 10 (for
 Wolf-Rayet stars. We also include the wind mass loss
 enhancement for fast rotating stars following \cite{maeder:00}.

We include the effects
of tides on the spins and orbital angular momentum using the
algorithm from \cite{hurley:02} based on the calculations of \cite{zahn:77,hut:81}.

When the radius of the most massive star exceeds the Roche radius,
(calculated using \citealt{eggleton:83} fitting formula)
we use the algorithm of \cite{claeys:14} to determine the
mass transfer rate (see their Eq.~10).

We limit the accretion rate of the
accretor to 10 times its Kelvin-Helmholtz timescale, i.e., we assume a
mass transfer efficiency $\beta_\mathrm{RLOF}\equiv |
\dot{M}_\mathrm{acc}|/|\dot{M}_\mathrm{don}|\equiv
\beta_\mathrm{th}$, with $\dot{M}_\mathrm{don}$ and
$\dot{M}_\mathrm{acc}$ temporal variation of the donor and accretor mass, respectively. Larger rates are
likely to drive the accretor out of thermal equilibrium\footnote{We do
not model the internal structure of the stars, therefore, we do not
follow its possible bloating during mass transfer, which could
potentially enhance its mass loss and related spin down.} and lead to
unstable mass transfer \citep[][]{neo:77,hurley:02}. However, for a
large portion of the parameter space, $\beta_\mathrm{th}$ results in a
rather conservative mass transfer \citep[][]{schneider:15}. Since the
efficiency of mass transfer is a major uncertainty in binary evolution
\citep[][]{demink:07}, we also consider the extreme parameter
variations with $\beta=0,\,0.5,$ and 1, respectively.

Matter that is not accreted is assumed to leave the system with
the specific angular momentum of the orbit of the accretor
\citep[e.g.,][]{soberman:97,vandenheuvel:17}, $h = \gamma_\mathrm{RLOF}
J_\mathrm{orb}/(M_\mathrm{don}+M_\mathrm{acc})$,
with $\gamma_\mathrm{RLOF}=M_\mathrm{don}/M_\mathrm{acc}$. In our
model variations, we also explore a scenario where the mass that is not
accreted leaves the binary from the outer Lagrangian point L2 and is assumed to form a circumbinary disk \citep[i.e.,
$\gamma_\mathrm{RLOF}=\sqrt{2}(M_\mathrm{don}+M_\mathrm{acc})^2/(M_\mathrm{acc}M_\mathrm{don})\equiv
\gamma_\mathrm{disk}$,][]{artymowicz:94}. We further consider the
assumptions that the mass that is not accreted leaves the system with the
orbital specific angular momentum, i.e.~$\gamma=1$, which is the standard model of \cite{belczynski:08, dominik:12, dominik:13}.

%qcrits
If, by the time the first star fills its Roche lobe, the mass
ratio of the accretor to the donor is smaller than a certain threshold
($M_\mathrm{acc}/M_\mathrm{don}<q_\mathrm{crit}$), we assume that the
system enters a common envelope phase. The threshold value
$q_\mathrm{crit}$ is uncertain and
depends on the evolutionary stage of the donor. In our fiducial simulation $q_\mathrm{crit,A}=0.65$ for a
main sequence (MS) donor \citep[][]{demink:07}, and $q_\mathrm{crit,B}=0.4$ for a Hertzsprung gap
donor \citep[][]{hurley:02}.
We also consider values of $q_\mathrm{crit,A}=0.25,\,0.8$ and $q_\mathrm{crit,B}=1,\,0.5,\,0$
in our model variations.

We treat common envelope evolution using the $\alpha_{\rm CE}
\lambda$-formalism \citep[][]{webbink:84, livio:88, dekool:90, hurley:02}. In all
variations, we use an analytic fit to the $\lambda_{g}$ values of
\cite{dewi:00} for the binding energy parameter $\lambda$. Those values do not include the energy stored in thermal
motions and the ionization state of the material within the
envelope. It is unknown what fraction of that internal energy is
useful in unbinding the envelope \citep[see, e.g.,][and references
therein]{dewi:00, ivanova:13}.  We assume $\alpha_{\rm CE} = 1$ for
our fiducial simulation, i.e., perfectly efficient use of the
liberated gravitational potential energy from the orbit but without
additional energy sources.  We explore variations with  $\alpha_{\rm
  CE} = 0.1$, i.e., inefficient use of the orbital energy to eject the
common envelope, and $\alpha_{\rm CE} = 10$, intended as indicative of
a fairly extreme case of additional energy input. For
binaries that interact while neither stars have a well defined core-envelope structure, such as
MS stars and helium-MS stars, we assume that the common envelope
phase leads to a merger as described in \cite{demink:13, schneider:15}.

\subsubsection{Natal kick}
\label{sec:SN_kick}

When a CC event happens in a binary system three different physical
ingredients contribute to possibly disrupting the system and ejecting the
companion star.

\begin{enumerate}
\item[i.] The orbit is modified by the sudden change in gravitational potential, because
of the mass lost through ejecta \citep[the so-called ``Blaauw
kick'',][]{zwicky:57,blaauw:61, boersma:61}. However, this is not the
dominant effect, because the typical evolution of a massive close binary
system involves mass transfer through Roche lobe overflow (RLOF) before the first CC (cf.~\Figref{fig:cartoon}). RLOF
removes the envelope of the initially more massive star, limiting
severely the amount of mass that can be ejected by the SN. The
``Blaauw kick'' alone rarely unbinds the system \citep[][]{huang:63,tutukov:73,leonard:94}.
\item[ii.] The SN shock can interact with the companion star.
  \citep[e.g.,][]{wheeler:75,liu:15, rimoldi:16, hirai:18}. The shock can dynamically
  remove mass from the companion (stripping), heat the envelope of the
  secondary, thus enhancing its own mass
  loss (ablation), and deposit mass and momentum on the secondary. The injection of energy can inflate the companion and make
  it look redder for a duration comparable to its thermal timescale. Our
  treatment of accretion only changes the mass of the companion, without
  checking for structural readjustments of the star. However, the
  interaction between the ejecta and the secondary is typically a small effect \citep[e.g.,][]{liu:15}.
\item[iii.]  The natal kick of the compact object changes %instantly
  its kinetic energy and momentum. This kick is caused by asymmetries
  in the SN ejecta and/or neutrino flux at the explosion, possibly
  seeded by the late core and shell burning phases of stellar
  evolution \citep[see e.g.,][]{shklovskii:70,kusenko:96,fryer:04,socrates:05,wongwathanarat:13,
    janka:13,janka:17,holland-ashford:17,grefenstette:17,katsuda:18}. It
is typically parametrized using a kick velocity $\mathbf{v}_k$, drawn
from a distribution of amplitude $v_k\equiv |\mathbf{v}_k|$ and directions (see below). The natal
kick is the dominant reason for the disruption of binaries.
\end{enumerate}

We have updated the treatment of binary disruptions by a CC event in
\texttt{binary\_c}, following \citealt{tauris:98} (TT98 hereafter). To re-calculate the post-CC
orbital parameters, their algorithm
assumes instantaneous
loss of the ejecta, because the ejecta velocity $v_\mathrm{ej}$ is much larger than the
orbital velocity $v_\mathrm{orb}$ ($v_\mathrm{orb} \lesssim 10^2\,\mathrm{km \
  s^{-1}} \ll v_\mathrm{ej}\sim10^4\,\mathrm{km \
  s^{-1}}$). This algorithm considers the motion of the compact
object within the pre-CC orbit, and the interaction between
the ejecta and the companion star. 
We set the
relative change of mass 
of the secondary due to stripping, ablation, and
accretion of SN ejecta following a fit to the three-dimensional
hydrodynamical simulations
of \cite{liu:15} for the impact of the ejecta on a
$M_2=3.5\,M_\odot$ star, which is the most massive companion they
considered. We also assume an efficiency of momentum
transfer\footnote{$\eta=0.5$ in the algorithm from \citetalias{tauris:98}.} from the
ejecta to the companion star of 0.5 (\citetalias{tauris:98};\citealt{hirai:18}). The contribution of 
the interactions between the SN ejecta and the secondary star to the total natal kick is
typically small, $\lesssim10\,\mathrm{km \ s^{-1}}$
(\citetalias{tauris:98}; \citealt{liu:15}; \citealt{hirai:18}).

>>>>

% \defcitealias{tauris:98}{TT98}
% \defcitealias{ramirez-agudelo:15}{R-A15}

@@

\title{VFTS682: a confirmed dynamical ejection?}

\author{M.~Renzo\inst{1} \and \todo{TBD}% S.~E.~de~Mink\inst{1} \and
  % S.~Justham\inst{2,3} \and A.~de~Koter\inst{1} \todo{etc..order to be decided}
  .}

\institute{{Astronomical Institute Anton Pannekoek, University of
    Amsterdam, 1098 XH Amsterdam, The Netherlands}
  \and{School of Astronomy \& Space Science, University of the Chinese
    Academy of Sciences, Beijing 100012, China}
  \and{National Astronomical Observatories, Chinese Academy of
    Sciences, Beijing 100012, China}
}

\offprints{M.~Renzo, \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}}
\date{}
\abstract{}

\keywords{stars: kinematics, stars:individual:VFTS682}
\maketitle{}

@@

How do stars form is one longstanding question in astrophysics
\todo{ref?}. Answering it for massive stars is particularly difficult,
since these are intrinsically rare
\citep[e.g.,][]{salpeter:55,kroupa:01, schneider:18}, evolve fast, and
remain enshrouded in their parent cloud during the whole formation
process. Moreover, observations of young massive stars reveal a
complicated multipliticy structure which requires explanation.

Understanding massive star formation, possibly as a function
of metallicity, is a key question given the present and upcoming
transient survey \citep[e.g., LSST, BlackGem, LIGO/Virgo O3][]{} which will reveal transients associated to their
evolution and death.

\todo{short summary of formation scenarios goes here}.

The second data release (DR2) from the Gaia satellite
\cite[][]{brown:18,} allows us to test
these hypothesis using one particular star, VFTS682. This star is a
very massive ($\sim150\,M_\odot$) WNh5 star in the 30 Doradus region
of the Large Magellanic Cloud, and it is presently observed at a
projected distance of $\sim$$29$\,pc from the nearest cluster of
massive stars R136 \citep[][]{bestenlehner:11}. Based on the extremely high mass of this star and
its present day apparent isolation, \cite{bestenlehner:11} proposed it
might be a candidate for isolated star formation, if it has not been
ejected from R136 in the past.

In this study, we combine the radial velocity measurements from the
VFTS survey \citep[][]{evans:11} with the proper motion from Gaia DR2
to reconstruct the three-dimensional velocity of VFTS682, and test the
null hypothesis that this star was ejected from R136. \todo{check the
  following} Our results indicate that R136 is the likely origin of
this star, and therefore isolated star formation is \emph{not}
required to explain it. However, we find a discrepancy between the
apparent age of VFTS682, it's kinematic age, and the age estimates for
R136.

In \Secref{sec:sample} we describe how we select stars in R136 and in
the surroundings of VFTS682 itself to remove the mean motion of the
region. \Secref{sec:results} presents our main findings. We conclude by
discussing the implications for theories of
star formation, N-body interactions, and binary evolution in
\Secref{sec:discussion}.

\section{Gaia DR2 data selection}
\label{sec:sample}
VFTS682 is labeled in the Gaia DR2
catalog\footnote{\url{https://gea.esac.esa.int/archive/}} with the
source id 4657685637907503744. For this star, we retrieve from the
Gaia catalog its position in right ascension (RA) and declination (DE)
in the IGCS frame\cite[][]{brown:18}\todo{some 2016 Gaia paper?}, its
proper motion components ($\mu_\mathrm{RA}$, and $\mu_\mathrm{DE}$,
respectively). For the radial velocity of VFTS16 and of the 30 Doradus
region as a whole, we instead use the VFTS data
as quoted in \cite{bestenlehner:11}. Our main conclusions do not
depend on the detailed value of the radial velocity.\todo{improve
  phrasing} The values of all these quantities that we use are listed
in \Tabref{tab:vfts682}.

\begin{table}[tbp]
  \centering
  \begin{tabular}[htbp]{l|c|c}
    Parameter & Value & Source\\ \hline\hline
    RA \hfill[degree] &  84.73 $\pm$  0.036 & \multirow{4}{*}{Gaia DR2}\\
    DE \hfill [degree] & -69.07 $\pm$  0.05  & \\
    $\mu_\mathrm{RA}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 1.84 $\pm$ 0.07 & \\
    $\mu_\mathrm{DE}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.78 $\pm$ 0.08& \\
    $\delta v_\mathrm{rad}$  \hfill[$\kms$] & -45 $\pm$ 25 & \cite{bestenlehner:11}\\
    \hline
  \end{tabular}
  \caption{Astrometric parameters for VFTS682}
  \label{tab:vfts682}
\end{table}

To compare the astrometry of VFTS682 and derive its peculiar motion,
we then select data from the Gaia DR2 catalog for two regions: the
``surroundings'' of VFTS 682, and the ``R136 cluster''. The
surrounding region is defined by all the stars within 10 arcminutes from
VFTS682. This includes the subset of stars which we consider to be
part of R136. We include all the stars for which a value of the proper
motion components and the corresponding errors are available.
The ``R136 cluster'' is effectively defined by taking all the stars
within 25 arcseconds from R136a, one of the most massive members of
the cluster itself \citep[][]{crowther:10}.

Throughout this study, we assume the same distance of $50$\,kpc to the star, and to
the 30 Doradus region as a whole, since the parallax for VFTS682
listed in the Gaia DR2 catalog is negative.

\section{The kinematics of VFTS682}

\begin{figure*}[htbp]
  \centering
  \includegraphics[width=\textwidth]{./figures/main_plot}  
  \caption{position and projected relative velocity to R136}
  \label{fig:main}
\end{figure*}

\bibliographystyle{aa}
\bibliography{bibliography/vfts682}
